Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-25 paper/high-assurance-contracts.tex
checkpointing

\subsubsection*{Wellformedness}
Stream-based evaluation allows values arbitrarily `deep in the past'
to be accessed by use of $\konst{pre}(-)$. However, the programming
languages we target do not directly support streams.  We are currently
developing a means of translating away such deep temporal accesses by
the introduction of intermediate variables. For now, we focus on the
notion of \emph{wellformed} equations that needs to hold. The guiding
intuition is that a wellformed {\Eqs} allows access at most one step
in the past, and accesses of a past value for a variable are only
possible until the variable is updated. In informal terms, this means
that ${\Eqs}\cdot G$ in a code contract must be such that:
\begin{itemize}
\item \konst{fby} only occurs at the top level of a variable definition, \eg, $v = e_1 \to e2$;
\item $\konst{pre}(e)$ is allowed only when $e$ is a variable; and
\item $\konst{pre}(v)$ is allowed before the defining equation $v = e$, and may occur in $e$.
\end{itemize}
For example, the Fibonacci sequence $1,1,2,3,5,8,13,\ldots$ can be
defined, albeit in a non-wellformed manner, by
{\small
\begin{verbatim}
  Fib = 1 -> pre(1 -> Fib + pre Fib)
\end{verbatim}
}
\noindent or by (again non-wellformed)
{\small
\begin{verbatim}
  N = 0 -> 1 + pre(N)
  Fib = if N <= 1 then 1 else pre(Fib) + pre(pre(Fib)).
\end{verbatim}
}
\noindent or by (wellformed):
{\small
\begin{verbatim}
  N = 0 -> 1 + pre(N)
  A = 42 -> pre(Fib)
  Fib = if N <= 1 then 1 else pre(Fib) + pre(A)
\end{verbatim}
}
A wellformed ${\Eqs} \cdot G$ is straightforward to implement in
conventional programming languages, as we will see in
Section \ref{sec:synthesis}.

@@

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \scalebox{0.62}{\usebox{\flt}} \\
    \end{tabular}
  \end{center}
  \caption{Code contract specification for high-assurance filter.}
  \label{fig:filter}
\end{figure}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
    \scalebox{0.62}{\usebox{\mntr}} \\
    \end{tabular}
  \end{center}
  \caption{Code contract specification for high-assurance monitor.}
  \label{fig:monitor}
\end{figure}

>>>>

\subsubsection*{Wellformedness}
Stream-based evaluation allows values arbitrarily `deep in the past'
to be accessed by use of $\konst{pre}(-)$. However, the programming
languages being targeted by SPLAT do not directly support streams.  We
are currently developing a means of translating away such deep
temporal accesses by the introduction of intermediate variables. For
now, we focus on the notion of \emph{wellformed} equations that needs
to hold. The guiding intuition is that a wellformed {\Eqs} allows
access at most one step in the past, and accesses of a past value for
a variable are only possible until the variable is updated. In
informal terms, this means that ${\Eqs}\cdot G$ in a code contract
must be such that:
\begin{itemize}
\item \konst{fby} only occurs at the top level of a variable definition, \eg, $v = e_1 \to e2$;
\item $\konst{pre}(e)$ is allowed only when $e$ is a variable; and
\item $\konst{pre}(v)$ is allowed only before the defining equation $v = e$ (and may occur in $e$).
\end{itemize}
For example, the Fibonacci sequence $1,1,2,3,5,8,13,\ldots$ can be
defined, albeit in a non-wellformed manner, by
{\small
\begin{verbatim}
  Fib = 1 -> pre(1 -> Fib + pre Fib)
\end{verbatim}
}
\noindent or by (again non-wellformed)
{\small
\begin{verbatim}
  N = 0 -> 1 + pre(N)
  Fib = if N <= 1 then 1 else pre(Fib) + pre(pre(Fib)).
\end{verbatim}
}
\noindent or by (wellformed):
{\small
\begin{verbatim}
  N = 0 -> 1 + pre(N)
  A = 42 -> pre(Fib)
  Fib = if N <= 1 then 1 else pre(Fib) + pre(A)
\end{verbatim}
}
A wellformed ${\Eqs} \cdot G$ is straightforward to implement in
conventional programming languages, as we will see in
Section \ref{sec:synthesis}.

@@

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \scalebox{0.62}{\usebox{\flt}} \\
    \end{tabular}
  \end{center}
  \caption{Code contract for high-assurance filter.}
  \label{fig:filter}
\end{figure}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
    \scalebox{0.62}{\usebox{\mntr}} \\
    \end{tabular}
  \end{center}
  \caption{Code contract for high-assurance monitor.}
  \label{fig:monitor}
\end{figure}

### 2022-03-25 paper/synthesis.tex
checkpointing

\emph{Discussion on mapping program from stream environment to logic
function and then to CakeML}

Synthesis maps from model and specifications to code. The synthesis
algorithm traverses the system architecture looking for occurrences of
filter and monitor specifications;  for each such occurrence it
generates a CakeML program. In the following, we examine both filter
and monitor synthesis. The latter is typically much more involved, and
we will therefore devote more attention to it.

>>>>

Synthesis maps from model and specifications to code. The SPLAT tool
traverses the system architecture looking for occurrences of high
assurance components specified by code contracts; for each such
occurrence it generates a CakeML program. Starting from the semantics
of the code contract described in Section \ref{sec:code-contracts}
a sequence of semantics-preserving steps are taken:
\begin{itemize}
\item $\sem{{\Eqs}\cdot G}$ yields a function on arbitrary depth streams
\item transform to function over depth-1 streams (wellformedness)
\item transform to logic function
\item attach buffer handling to inputs and outputs (contig type parsing)
\item translate to CakeML
\item run CakeML compiler
\end{itemize}

In the following, we examine further details in both filter and
monitor synthesis.
